Response to reviewers

### 2022-05-04 paper/paper.tex
Adress minor comments.

\begin{abstract}
Textbook data is essential for teaching statistics and data science methods because they are clean, allowing the instructor to focus on methodology. Ideally textbook data sets are refreshed regularly, especially when they are subsets taken from an on-going data collection. It is also important to use contemporary data for teaching, to imbue the sense that the methodology is relevant today. This paper describes the trials and tribulations of refreshing a textbook data set on wages, extracted from the National Longitudinal Survey of Youth (NLSY79) in the early 1990s. The data is useful for teaching modeling and exploratory analysis of longitudinal data. Subsets of NLSY79, including the wages data, can be found in supplementary files from numerous textbooks and research articles. The NLSY79 database has been continuously updated through to 2018, so new records are available. Here we describe our journey to re-create the wages data, and document the process so that the data can be regularly updated into the future. Our journey was difficult because the steps and decisions taken to get from the raw data to the wages textbook subset, have not been clearly articulated. We have been diligent to provide a reproducible workflow for others to follow, which also hopefully inspires more attempts at refreshing data for teaching. Three new data sets and the code to produce them are provided in the open source R package, called {[}CENSORED{]}.
\end{abstract}

@@

For \texttt{gender} and \texttt{race}, we only rename the column names provided in the raw data, \texttt{new\_data\_qnames}. It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commonly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female''. The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similarly, this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

@@

\begin{itemize}
\tightlist
\item
  If the hourly rate is recorded as 0, we set wage as missing.
\item
  If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hour worked as missing.
\end{itemize}

@@

For \texttt{stwork} variable, we only rename the column names from \texttt{new\_data\_qnames}. This variable is then used to calculate the next variable, \texttt{yr\_wforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual started working (\texttt{stwork}). Finally, \texttt{exp} variable is derived from the number of weeks worked since the last interview indicated as variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

In the next step, we explore the mean hourly wage data of samples of individuals. The purpose is to examine the common patterns and check the quality. A random sample of 36 individuals is chosen (using the \texttt{sample\_n\_keys} function in \texttt{brolgar}). Their longitudinal profiles are plotted, faceted by \texttt{id}, and using free \(y\) scales so that the individual patterns can be examined (Figure \ref{fig:sample-plot}). There is a lot of variability from one individual to another and substantial fluctuation in wages at different times for most individuals. Some individuals (2799, 11041, 11146) are only measured for a short period. Some individuals (8296, 9962) possibly have errors in wages in some years because of the extreme fluctuation. These need to be inspected more closely. It is also important to note that some shorter profiles indicate that some individuals have left the study before it has finished. Checking whether the demographics of the early departing are similar to those who remain in the study is an important part of any downstream analysis to account for the bias induced by the inadvertent censoring.

Figure \ref{fig:feature-plot}s shows an alternative way to check the data quality. Plot (A) is the spaghetti plot where all profiles are shown, and it can be seen that there are unbelievably high wage values (up to \$60,000/hour) for some individuals, mostly around 1990. Plot (B) shows side-by-side boxplots of the three number summaries (minimum, median, and maximum) for all individuals. This tells us that there are a number of individuals with unbelievably high maximum wages. Plot (C) shows the profile for an individual, with not such a high maximum wage but still indicates a problem: their wages are consistently low except for one year where they earned close to \$1200/hour. This does not seem to be reasonable and leads us to use a procedure to detect and fix these temporal anomalies.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate, that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analysts could filter them later.
\item
  79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because:

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    35 of them completed at least 12th grade with a diploma instead of GED (\texttt{ged} variable is coded to 1). This suggests that they are not dropouts, and so we excluded them from the refreshed data.
  \item
    The information about \texttt{ged} is missing in 38 individuals. We decided to include them in the refreshed data.
  \item
    3 individuals have both diploma and GED (\texttt{ged} is coded to 3). These were kept in the refreshed data.
  \item
    12 individuals do not exist in \texttt{wages} data because they have participated in less than 3 rounds of the survey.
  \end{enumerate}
\end{enumerate}

@@

There are several aspects of the original data that were difficult to replicate. The calculation of the work experience is not clearly articulated. Singer and Willett (2003) describe the temporal variable as the years of experience since the first day of work. This variable is not explicitly available in the database. From the NLSY79 topical guide (Bureau of Labor Statistics, U.S. Department of Labor 2021b), we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. This is used to calculate the variable. It produces reasonably similar but not exactly the same values. Because the original data set did not include the year of the survey, it cannot be precisely compared.

The highest grade completed has some confusion. There are several ways that this is reported, including the highest grade ever completed and also the \texttt{hgc} at each survey year. To match the original data, it is appropriate to use the \texttt{hgc} while in high school. The documentation suggests that this variable is available, but it is not actually present in the data. Hence, for matching the original data, we have calculated the \texttt{hgc} to match the \texttt{hgc} achieved in the years between 1979 and 1994 based on the yearly survey value. The result does not exactly match with the original for a few individuals.

@@

For the data providers, we recommend that a better validation system with clear rules applied at data entry and that alternative output formats, such as a tidy format, would help users make better use of their resources. The problem with many of the wages records is that there are implausible values or confusion on how to record wages for multiple jobs. These values can be validated with simple checks at data entry. Providing an open data resource also is accompanied by the responsibility that the data, especially data that is as valuable as this, is reliable. Users need to be able to trust the data.

@@

The wages data provides a good opportunity to discuss the difference between statistics to use for public policy and statistics that relate to the individual. Public policy is based on models, yielding averages that might vary across strata. Modeling the wages relative to workforce experience, with demographic covariates, we learn that there are significantly different patterns. That more education leads to increasingly higher wages, which is a satisfying result for educators. It means that education makes a difference in the wage experience and that public policy that encourages education is a data-supported action. We would also learn, although not presented here, that race matters and that being black leads to lower wages. This is a disturbing finding because there is no rationale for such a difference in a fair society. This would provide support for action in public policy to remove this overall average effect. It also provides an example for educators to explain the use of data to support public policy action.

On an individual level, one needs to know where I am in this data and does this data relate to me. To do this, the individual profiles need to be explored. Pre-dominantly, we would learn that the variation from one individual to another is far more than the variation between demographic strata. For example, many individuals with lower educational attainment earn very high wages. From a statistics and data science educator perspective, more focus and more methodology for this type of statistics need to be included in the curriculum.

The above interpretations of results from analyzing the wages data rely on trusting that the data provided is accurate and valid. The wages data is collected by a reputable organization, but we found that the data has obvious errors that should be corrected. This paper has illustrated procedures and guidelines to achieve valid data and provides the code and details for it to be reproduced and modified if deemed appropriate.

>>>>

\begin{abstract}
Textbook data is essential for teaching statistics and data science methods because they are clean, allowing the instructor to focus on methodology. Ideally textbook data sets are refreshed regularly, especially when they are subsets taken from an on-going data collection. It is also important to use contemporary data for teaching, to imbue the sense that the methodology is relevant today. This paper describes the trials and tribulations of refreshing a textbook data set on wages, extracted from the National Longitudinal Survey of Youth (NLSY79) in the early 1990s. The data is useful for teaching modeling and exploratory analysis of longitudinal data. Subsets of NLSY79, including the wages data, can be found in supplementary files from numerous textbooks and research articles. The NLSY79 database has been continuously updated through to 2018, so new records are available. Here we describe our journey to re-create the wages data, and document the process so that the data can be regularly updated into the future. Our journey was difficult because the steps and decisions taken to get from the raw data to the wages textbook subset have not been clearly articulated. We have been diligent to provide a reproducible workflow for others to follow, which also hopefully inspires more attempts at refreshing data for teaching. Three new data sets and the code to produce them are provided in the open source R package called {[}CENSORED{]}.
\end{abstract}

@@

For \texttt{gender} and \texttt{race}, we only rename the column names provided in the raw data, \texttt{new\_data\_qnames}. It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commonly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female.'' The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustments of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similarly, this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

@@

\begin{itemize}
\tightlist
\item
  If the hourly rate is recorded as 0, we set wage as missing.
\item
  If the total hours of worked for the corresponding job is greater than 84 hours, we set the wage and hours worked as missing.
\end{itemize}

@@

For \texttt{stwork} variable, we only rename the column names from \texttt{new\_data\_qnames}. This variable is then used to calculate the next variable, \texttt{yr\_wforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual started working (\texttt{stwork}). Finally, \texttt{exp} variable is derived from the number of weeks worked since the last interview indicated as variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is weeks, we convert this to year.

@@

In the next step, we explore the mean hourly wage data of samples of individuals. The purpose is to examine the common patterns and check the quality. A random sample of 36 individuals is chosen (using the \texttt{sample\_n\_keys} function in \texttt{brolgar}). Their longitudinal profiles are plotted, faceted by \texttt{id}, and using free \(y\) scales so that the individual patterns can be examined (Figure \ref{fig:sample-plot}). There is a lot of variability from one individual to another and substantial fluctuation in wages at different times for most individuals. Some individuals (2799, 11041, 11146) are only measured for a short period. Some individuals (8296, 9962) possibly have errors in wages in some years because of the extreme fluctuation. These need to be inspected more closely. It is also important to note that some shorter profiles indicate some individuals have left the study before it has finished. Checking whether the demographics of the early departing are similar to those who remain in the study is an important part of any downstream analysis to account for the bias induced by the inadvertent censoring.

Figure \ref{fig:feature-plot} shows an alternative way to check the data quality. Plot (A) is the spaghetti plot where all profiles are shown, and it can be seen that there are unbelievably high wage values (up to \$60,000/hour) for some individuals, mostly around 1990. Plot (B) shows side-by-side boxplots of the three number summaries (minimum, median, and maximum) for all individuals. This tells us there are a number of individuals with unbelievably high maximum wages. Plot (C) shows the profile for an individual with a maximum wage that is not so extreme but still indicates a problem: their wages are consistently low except for one year where they earned close to \$1200/hour. This does not seem to be reasonable and leads us to use a procedure to detect and fix these temporal anomalies.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analysts could filter them later.
\item
  79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because:

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    35 of them completed at least 12th grade with a diploma instead of GED (\texttt{ged} variable is coded to 1). This suggests they are not dropouts, and so we excluded them from the refreshed data.
  \item
    The information about \texttt{ged} is missing in 38 individuals. We decided to include them in the refreshed data.
  \item
    3 individuals have both diploma and GED (\texttt{ged} is coded to 3). These were kept in the refreshed data.
  \item
    12 individuals do not exist in \texttt{wages} data because they have participated in less than 3 rounds of the survey.
  \end{enumerate}
\end{enumerate}

@@

There are several aspects of the original data that were difficult to replicate. The calculation of the work experience is not clearly articulated. Singer and Willett (2003) describe the temporal variable as the years of experience since the first day of work. This variable is not explicitly available in the database. From the NLSY79 topical guide (Bureau of Labor Statistics, U.S. Department of Labor 2021b), we find several variables are tagged as work experience-related variables. One of them is the weeks worked since the last interview. This is used to calculate the variable. It produces reasonably similar but not exactly the same values. Because the original data set did not include the year of the survey, it cannot be precisely compared.

The highest grade completed has some confusion. There are several ways this is reported, including the highest grade ever completed and also the \texttt{hgc} at each survey year. To match the original data, it is appropriate to use the \texttt{hgc} while in high school. The documentation suggests this variable is available, but it is not actually present in the data. Hence, to match the original data, we have calculated the \texttt{hgc} to match the \texttt{hgc} achieved in the years between 1979 and 1994 based on the yearly survey value. The result does not exactly match the original for a few individuals.

@@

For the data providers, we recommend a better validation system with clear rules applied at data entry and alternative output formats, such as a tidy format, which would help users better use their resources. The problem with many wages records is that there are implausible values or confusion on how to record wages for multiple jobs. These values can be validated with simple checks at data entry. Providing an open data resource is also accompanied by the responsibility that the data, especially as valuable as this, is reliable. Users need to be able to trust the data.

@@

The wages data provides a good opportunity to discuss the difference between statistics for public policy and statistics that relate to the individual. Public policy is based on models, yielding averages that might vary across strata. Modeling the wages relative to workforce experience with demographic covariates, we learn that there are significantly different patterns. More education leads to increasingly higher wages, which is a satisfying result for educators. It means education makes a difference in the wage experience and that public policy that encourages education is a data-supported action. We would also learn, although not presented here, that race matters and being black leads to lower wages. This is a disturbing finding because there is no rationale for such a difference in a fair society. This would provide support for action in public policy to remove this overall average effect. It also provides an example for educators to explain the use of data to support public policy action.

On an individual level, people might want to see how their characteristics relate to those in the dataset. To do this, the individual profiles need to be explored. Predominantly, we would learn the variation from one individual to another is far more than the variation between demographic strata. For example, many individuals with lower educational attainment earn very high wages. From a statistics and data science educator perspective, more focus and more methodology for this type of statistics need to be included in the curriculum.

The above interpretations of results from analyzing the wages data rely on trusting that the data provided is accurate and valid. The wages data is collected by a reputable organization, but we found the data has obvious errors that should be corrected. This paper has illustrated procedures and guidelines to achieve valid data and provides the code and details for it to be reproduced and modified if deemed appropriate.

### 2022-05-04 response/response2.Rmd
Adress minor comments.

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

* Yes, it is cut off, we have made the size smaller, so it is not cut off.

@@

* These have been done in the paper so we have removed it.

@@

* We have changed the word "mutate" to be "create".

@@

* We have removed unnecessary instances.

@@

* All of the extraneous punctuation marks mentioned have been removed.

@@

* We have added the missing closing parenthesis.

@@

* Done.

@@

* All of the comments on the list above have been addressed. (**Except for the Cooksey reference**)

@@

* All of the comments on the list above have been addressed.

@@

* Yes, you are correct, we have edited the point 3 to improve the clarity of this sentence.
